PhD Title \\ PhD Thesis

### 2024-07-31 Latex_files/Thesis/chapters/chapter 4/main.tex
corrections made to chapters 2 and 4 of thesis doc

In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Positional and Orientational dependence of Trapping forces}

If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus. The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser. We can do so because the optical force is only dependent on the particle's relative position. If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.

\begin{figure}[h!]
	\centering
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{.475\linewidth}
		\centering
		\raisebox{60pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
		\label{large over small}
	\end{subfigure}
	\begin{subfigure}{.475\linewidth}
		\centering
		\includegraphics[width=0.3\linewidth,
		keepaspectratio]{theta=180.png}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2. (a) is the case where the smaller sphere is orientated with the beam propagation direction. (b) is the inverted case, smaller sphere oriented against the propagation direction. Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
\end{figure}

We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable. As shown below in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased. This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap. This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\label{fig:transverse_force}
	\includegraphics[width=\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane. With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
\end{figure}

For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap. As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength. However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus. Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}. Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}. This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions. (a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position. (b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
\end{figure}

\subsection{Non-trivial harmonic traps}
\label{sec:off-axis}
Computing the equilibrium positions when a dimer is aligned with the electric field is relatively simple as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap. However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}. Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations. An example for a dimer of size ratio 2 is shown below:
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.67\linewidth}
		\includegraphics[width=\linewidth]{off_axis_trap.png}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
		\caption{}
	\end{subfigure}
	\captionsetup{hangindent=2.35cm}
	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW. The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in. Right hand render shows a dimer in its off-axis configuration}
\end{figure}

Interestingly while the trap strength of these off-axis traps are 
similar in magnitude to the vertically aligned dimers, but when 
the laser power is lowered (around 5 mW) the traps become metastable 
resulting in the dimer escaping from after some random time while 
within the trap. If the overall potential depth could be characterised 
then dimers placed into this orientation could be used as a 
micro-scale temperature alarm, where by fine tuning of the dimer's 
parameters would allow you to construct a potential well that can 
only be escaped when the local fluid temperature exceeds a certain maximum value.

@@

One aspect that has yet to be covered in depth with regards to
spherical aggregates of any construction is their interaction with
circularly polarised light. While most homogenous objects are trapped
no differently in different polarisations, dimers have been shown to
experience an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018}. 
Typically the spin density of an electric field cannot be reduced in 
homogenous medium due to the fact that the spin angular momentum is 
conserved locally. In which case the only way to transfer angular 
momentum to use an object that is anisotropic. This is seen most 
clearly with birefringent materials, due to fact that the crystal 
lattice itself is anisotropic, a circularly polarised beam will 
transfer an optical torque that is proportional to the difference in 
the refractive index of the particle. The optical torque is given as:
\begin{equation}
	\label{eq:opt_torque}
	\begin{aligned}
		\tau_{opt} = \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
	\end{aligned}
\end{equation}

Where $\Delta n$ is the difference in refractive indices of the 
particle, and $\phi$ indicates the phase shift in the electric field
(for circular polarised light $\phi= \pi/4$). The optical rotation seen
with dimers can be attributed to the polarisation of the trapping beam as
taking a Fourier transform of the transverse components of the dimers 
orientation time series reveals a clean rotation rate. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.9\linewidth]{rotation_vs_pol.png}
	\caption{Rotation frequency vs component phase difference for differently sized dimers. The solid lines represent the rotation rate experienced while the dimer is in its standard orientation, whereas the solid points are for the case where the orientation is inverted. Laser power $= 100\ mW$}
\end{figure}

A possible explanation of this phenomena was devised by by \cite{Yevick2017}, 
who found that highly focused Gaussian beams could produce second 
order effects in the Rayleigh regime resulting in a photo-kinetic 
force that results in orbital motion about the beam's central axis. 
This effect is rather minimal for single sphere's, resulting in a 
orbital frequency on the order of $10^{-1}\,{\rm Hz}$, with an 
order of magnitude difference when trapping aggregates of spheres. 
They computed the circulation rate by computing the time-average 
probability flux; however, when extended to the Mie regime we 
see a completely different behaviour, instead experiencing an 
optical torque about their long axis.S This rotation was first 
noted by Vigilante and co-workers who only considered this
behaviour for a symmetric dimer \cite{Vigilante2020}; we run number of
simulations for differently sized dimers in a circularly polarised
trap and looked at the rotation rate.  We found that not only is the
rotation rate dependent on the size of the dimer, but also on its
orientation and therefore their axial position.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
  \caption{Rotation rate plotted against dimer size ratio for a variety of different simulation scenarios. The red line is for the case where the larger sphere is above the smaller sphere. The blue line is the inverted case, while the initial position is above the focus of the trap. And lastly the green line is again for the inverted case, but when the dimer's initial position is below the focus of the trap.}
\end{figure}

It is difficult to see from the graph, but the rotation rate never
truly goes down to zero, reaching a minimum of $2\,{\rm Hz}$, which would
imply that a second sphere of radius $200\,{\rm nm}$ is enough to
induce rotational motion. We used \textit{mstm} to look at the stokes
parameters from the scattered field from a simple plane wave incident
on our dimer, the proportion of circularly polarised light is minimal
compared to the proportion of plane polarised light, which indicates
that this rotational motion is not due to any inhomogeneity in the
dimer that might impart angular momentum to the scattered beam - as
compared to a anisotropic scatterer like vaterite.

These results are somewhat contrary to other work with silica dimers
\cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments
have trapped the dimer in an orientation perpendicular to the beam
propagation direction.  The rotational motion is induced due to the
asymmetric geometry creating an unbalanced polarisation susceptibility
along its long axis as compared to its short axis; therefore its long
axis is aligned with the polarisation vector and can rotate
freely\cite{Ahn2018}.  This however cannot be the case with our
simulations as the dimer rotates about its long axis, meaning there
cannot be an asymmetric axis to align with the beam's polarisation
vector. Furthermore, we see a non-linear increase in the rotational
speed of our dimers with size, the drag torque from the surrounding
fluid is $\propto r^3$ so the expectation is that the rotation
frequency should fall off with increasing size.  This indicates that
the rotational motion is due to the shape asymmetry of the dimer and
not solely due to the beam's angular momentum.  Measurement of this
photo-kinetic force is difficult to achieve due to the fact that
previous analysis was conducted in the Rayleigh regime, where the
polarizability of our dimer can be approximated as:
\begin{align}
  \bold{p}({\bf r},t)
  =
  \alpha_x E_x({\bf r},t)\hat{\bf e}_x
  + \alpha_yE_y({\bf r},t)\hat{\bf e}_y
  + \alpha_zE_z({\bf r},t)\hat{\bf e}_z
\end{align}

@@

\subsection{Gyroscopic Precession using asymmetric dimers}
As mentioned in section~\ref{sec:off-axis} for specificity 
sized dimers there is the potential for non-vertical trapping 
orientations in which the dimer is located in a harmonic 
trap. When trapped by a circularly polarised light these 
dimers exhibit gyroscopic precession. As shown in fig~\ref{fig:gyro} 
the dimer's is rotating about it's long axis (as seen by 
the periodic behaviour of $u_x$ and $u_y$), while also precessing 
around the optical axis of the beam, resulting in a second 
order oscillation in $s_z$.
\begin{figure}[h]
	\label{fig:gyro}
	\centering
	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
	\caption{3 second trajectory of a dimer trapped in an off axis orientation, laser power was set to 100 mW. The far left column depicts the dimer's centre of mass position with time; middle two columns are the $x$, $y$, and $z$ components of the vectors $u_x$ and $u_y$; last column depicts the components of the vector $u_z$ which defines the dimer's orientation.}
\end{figure}

@@

The linear relationship could partly be due to fact that we do not
account for the change in viscous forces with increasing
laser power. It is far more likely that the rotational frequency
reaches a maximum value assuming that the bulk fluid can readily 
absorb the laser.

This gyroscopic motion has been demonstrated previously 
in nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} 
but has not been observed for micron scale aggregates. 
Since the torque is computed by evaluating the beam coefficients 
it is be difficult to apply this result to micro-rheology 
experiments as one would need to know the exact magnitude 
of the optical torque ahead of time in order to make 
estimations about the local fluid viscosity. This is 
trivial for a birefringent spherical particle, less 
so for spherical aggregates whose equilibrium position and 
orientation are unknown.

>>>>

With the initial goal of the PhD being to induce nucleation events via a 
spherical micro-rotor the goal of this chapter was to - in a limited capacity
simulate and investigate the influence of a second particle being bound to our 
target sphere. The choice of a dimer, instead of an amorphous solid that 
might better represent a growing crystalline solid, allows us to consider 
how the dynamics of the aggregate change by varying the size factor. We build 
upon the works of Vigilante \textit{et al} \cite{Vigilante2020} to consider 
asymmetric dimers and how varying size parameters alters the dynamics and 
additionally makes characterising their interactions within an optical trap 
more cumbersome. Attempting to simulate an amorphous aggregate is rather
difficult as calculating the optical force and torque is computationally 
slow and orientation specific.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Positional and Orientational dependence of Trapping forces}
\label{sec:eq_positions}
If we wanted to start from first principles and determine the trap strength 
on our target particle the first step would be to locate the equilibrium position
relative to the trap focus. For a single sphere it is easy to enough to understand 
that its centre of mass will be drawn to focal point of the laser due to gradient
forces, once there the force is analogous to a harmonic spring with a fixed trap
stiffness. Now, if we consider instead a dimer, we now have two spheres both being
drawn to the focus along by the same gradient force; in addition the scattering force
is significantly more complex due to both spheres scattering the electromagnetic 
fields. This mutual scattering between individual spheres is what makes simulating
spherical aggregates far more difficult compared to a single sphere, and even harder
still to predict the position where the dimer's centre of mass is in equilibrium.

Because the scattering force is only significant in the direction of beam 
propagation the potential well in the transverse plane can still be assumed 
to be harmonic around the central beam axis. The axial optical force cannot 
be assumed to behave as a simple harmonic trap. The methodology for computing 
optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the 
trapping force and determine where a simple sphere would be located relative 
to focal point of the laser by finding the position that minimises the net 
optical force in a negative feed back loop ($\delta F/\delta x < 0$) - we can 
assume that for a dielectric sphere the optical torque is negligible. For a 
dimer (or any arbitrary spherical aggregate), we now must consider both its 
position and orientation and find where the net optical force and torque are 
minimised.

\paragraph{Simulation Parameters}
As a paradigmatic example, consider a dimer suspended in water ($n_p = 1.59, 
n_m = 1.33$) located at the focus of a Gaussian beam (more specifically a Laguerre-Gaussian beam of mode $[0,0]$), the beam is focused by a objective 
with numerical aperture of 1.2 and is x polarised. The size ratio of the two 
sphere's is given by $a_{I}/a_{II} = 2$ where $a_{I}$ is kept at $1\ \mu m$ 
unless specified otherwise; the dimer's orientation is given by a unit vector connecting the centres of both spheres, we define the 'standard' orientation 
as being aligned with the direction of beam propagation direction - and 
therefore the 'inverted' orientation is defined when the dimer is orientated 
against the direction of beam orientation (see figures~\ref{fig:paradigmatic}(b) 
and (d)).

After computing its $T$-matrix via \textit{mstm} and supplying that 
to \textit{ott} we compute the optical force via \eqref{eq:optical_force} 
in the axial direction while the dimer is in its 'standard' orientation. 
As expected we see a single point where the dimer will be in equilibrium, 
the linear fit in fig.~\ref{fig:paradigmatic}(a) indicates a that the force 
can be modelled as a harmonic potential close to the equilibrium position ($F_z\approx-\kappa_z z$). The second point where the axial 
force goes to 0 cannot be considered as equilibrium position as the positive 
gradient indicates that the trap is unstable unless Brownian motion is ignored.

We repeated the same calculation but now while the dimer is in its 'inverted' orientation, instead of a single point where the optical force is minimised we 
see that there are instead two separate equilibrium positions, one above the focus
and one below the focus. In this particular example the two positions are far enough
apart that both can be considered as separate harmonic traps.    
\begin{figure}[h!]
	\centering
	\begin{subfigure}{.75\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{50pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
		\caption{}
		\label{large over small}
	\end{subfigure}
	\medskip
	\begin{subfigure}{.75\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{30pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
		\caption{}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the centre of mass of 
		the dimer (µm) for the case of a dimer of size ratio 2. (a) is 
		the case where the dimer is in its' 'standard' orientation. 
		(c) is the case where the dimer is in its' 'inverted' orientation 
		(b) and (d) are renders to visualise the dimer orientation are shown 
		below each plot. The black lines on each force-curve is a linear 
		fit with the slope being reported as the trap stiffness in the legend.}
		\label{fig:paradigmatic}
\end{figure}

We can see that both equilibrium positions have comparable axial trap stiffness 
($\kappa_z$), however the difference in the transverse trap stiffness ($\kappa_x$) 
is far more noticeable. As shown in Fig~\ref{fig:transverse_force}, 
the dimer's orientation and relative position significantly changes the force 
curve; not only is the trap wider when inverted but the trap stiffness is 
increased. This highlights one of the challenges involved with studying 
asymmetric particles, even though its a simple enough process to trap them 
they maybe characterised very differently depending on their relative position 
and orientation towards the focus. This can have a significant impact on 
rheological studies - or attempting to probe any local property - as the 
variance in trap strength can result in large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the dimer's centre of mass spheres, 
		where a positive force indicates the dimer is directed right on the x-axis, 
		and vice versa for a negative force. The same simulation parameters are used here as in fig~\ref{fig:paradigmatic}(a) and (c). The blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
	\label{fig:transverse_force}
\end{figure}

\newpage
For completeness the harmonic traps were located for dimers across a range 
of size ratios - from $a_{I}/a_{II} = 1$ to $a_{I}/a_{II}=10$ - while also 
recording the trap stiffness for each trap. The same simulation parameters are 
used here as for figures \ref{fig:paradigmatic} \& \ref{fig:transverse_force}. 
As shown in Fig.~\ref{fig:eq_positions} $a_{II}$ decrease the dimer begins to
approximate a single homogenous sphere - at least in terms of location and trap strength. However, for intermediate sized dimers (between $a_{I}/a_{II} = 1.1$ 
to $a_{I}/a_{II}=4$), a second equilibrium position is found below the trapping 
focus. Previous work using the ray-optics model have confirmed even in the case 
that two spheres begin separated the electric field will align the particles as 
such that they make contact and are trapped together about a single trapping 
position \cite{Xu2005}. Furthermore it has been shown through proper manipulation 
of the Gaussian or Bessel beam modes that any number of trapping potentials can 
be developed \cite{Shahabadi2020} for nanoparticles. This result however, is the 
first example of an orientation dependent trapping situation using only a 
$TEM_00$ beam. Typical experimental arrangements cannot determine much information 
on the axial position of a trapped particle relative to the trap focus; this 
result indicates not only that dimers can be trapped in multiple axial positions 
but also their trapping behaviour is heavily dependent on said axial position. 
As such it is necessary that positional information in the z-axis can be elucidated
if multiple spheres are trapped simultaneously.

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying 
		size ratio, dashed lines represent unstable traps whereas solid lines 
		are for stable equilibrium positions. (a) shows that dimers while in their
		'standard' orientation will always have a single equilibrium position. 
		(b) shows that when the same dimer is in its' 'inverted' orientation 
		can be trapped in two axial positions, one below the focus and one above 
		the focus.}
	\label{fig:eq_positions}
\end{figure}

\newpage
\subsection{Non-trivial equilibrium configurations}
\label{sec:off-axis}
Computing the equilibrium positions when a dimer is aligned with the electric 
field is relatively simple as the orientational torque is minimised (see
Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change
orientation enough to escape the trap. However, that does not rule out the 
possibility that there is a stable configuration where the orientation not 
strictly vertical, in fact most experimental work with symmetric nano-dimers 
will trap them lying perpendicular to the beam direction \cite{Ahn2018, 
Reimann2018}. Unlike in Sec.~\ref{sec:eq_positions} we cannot simply measure 
the optical force and torque as the parameter space is too large and determining 
if a particular position and orientation is stable is not clear based solely 
on force and torque measurements \cite{Bui2017}. Using the same simulation 
parameters as before we ran a number short simulations (total simulation time 
was $0.005$ s) with the laser power increased to $500\ mW$. Each simulation 
started with the dimer in a different starting position and orientation, due 
to the high laser power the dimers either escaped the trap or were stably 
trapped. The $z-\theta$ phase space - where $\theta$ is the angle between the 
direction of beam propagation and the dimer's orientation vector ($\theta=0^
\circ$ is the 'standard' orientation) - can be divided into different regions 
depending on which equilibrium configuration is reached.

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.67\linewidth}
		\includegraphics[width=\linewidth]{off_axis_trap.png}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
		\caption{}
	\end{subfigure}
	%\captionsetup{hangindent=2.35cm}
	\caption{Map of $z-\theta$ phase space using a dimer of size ratio 2 with 
		a laser power of 500 mW ($\theta=0\circ$ is the 'standard' orientation 
		and $\theta=180^\circ$ is the 'inverted' orientation). The stable 
		configurations are indicated by the larger circles and the starting 
		conditions are colour coded to match the stable point they end up in. 
		Right hand render shows a dimer in its off-axis configuration.}
	\label{fig:off_axis}
\end{figure}

Interestingly while the trap strength of these off-axis traps are similar 
in magnitude to the vertically aligned dimers, but when the laser power 
is lowered (around $5\ mW$) the traps become metastable; after reaching its'
equilibrium configuration the particle behaves similarly to a typically
trapped dimer but the trapping potential is small enough that the dimer
escapes in as little time as less than a second or after nearly a full 
3 seconds. Running similar simulations but for vertical configurations 
sees the dimer remaining trapped, even after 30 seconds of run time, 
indicating that the trapping potential is far greater than the thermal 
energy. This suggests that the reason this off-axis configuration is due 
to the rotational motion more than translational motion. If the overall 
potential depth could be characterised then dimers placed into this 
orientation could be used as a micro-scale temperature alarm, where 
by fine tuning of the dimer's parameters would allow you to construct a 
potential well that can only be escaped when the local fluid temperature 
exceeds a certain maximum value.

@@

One aspect that has yet to be covered in depth with regards to
spherical aggregates of any construction is their interaction with
circularly polarised light. While most homogenous objects are trapped
no differently in different polarisations, dimers have been shown to
experience an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018}. 
Typically the spin density of an electric field cannot be reduced in 
homogenous medium due to the fact that the spin angular momentum is 
conserved locally. In which case the only way to transfer angular 
momentum to use an object that is anisotropic. This is seen most 
clearly with birefringent materials, due to fact that the crystal 
lattice itself is anisotropic, a circularly polarised beam will 
transfer an optical torque that is proportional to the difference in 
the refractive index of the particle's ordinary and extraordinary axis. 
The total optical torque is given as:
\begin{equation}
	\label{eq:opt_torque}
	\begin{aligned}
		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
	\end{aligned}
\end{equation}

Where $\Delta n$ is the difference in refractive indices of the 
particle, $\phi$ indicates the phase shift in the electric field
(for circular polarised light $\phi= \pi/4$), and $\theta$ is the angle
between the particle's main axis and the direction of the electric field.
The optical torque of non-birefringent spherical aggregates has no such 
relationship as the mechanism behind their rotation is not as clearly defined.

To visualise the relationship between the polarisation state of the 
trapping beam, and the rotation rate we simulated the motion of an 
optically trapped dimer in beams of varying polarisation ($NA=1.2$, 
$P=100\ mW$), and the dimer is composed of silica ($n_p=1.59$, 
$n_m=1.33$). Each simulation was run for $1$ second ($\Delta t 
=10^-5$) and at the end we looked at the orientational time series; 
the dimer's orientation is recorded as a quaternion which can be easily 
converted to a 3-dimensional rotation matrix. By considering only the 
transverse components ($U_{1,x}$, $U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) 
of the rotation matrix and taking the Fourier transformation of their 
time series reveals the rotational frequency. The laser power is set to 
$100\ mW$ to avoid large rotational fluctuations and so that the Fourier 
series of the transverse components approximates $\delta(\omega_{rot}-f)$ 
- the Dirac delta function centred at the rotational frequency $\omega_{rot}$.

%% Example plot of rotating beam trajectory and then the same thing but 
%% Fourier transformed
\begin{equation}
\begin{split}
	q(t) \rightarrow R(t) =& 
	\begin{pmatrix}
		U_{1,x}(t) && U_{2,x}(t) && U_{3,x}(t) \\
		U_{1,y}(t) && U_{2,y}(t) && U_{3,y}(t) \\
		U_{1,z}(t) && U_{2,z}(t) && U_{3,z}(t) 
	\end{pmatrix} \\
	\rightarrow
	&\int^\infty_{-\infty}R(t)e^{-i2\pi ft} dt = 
	\begin{pmatrix}
		\delta(\omega_{rot}-f) && \delta(\omega_{rot}-f) && \delta(f) \\
		\delta(\omega_{rot}-f) && \delta(\omega_{rot}-f) && \delta(f)\\
		\delta(f) && \delta(f) && \delta(f)
	\end{pmatrix}
\end{split}
\end{equation}

If the rotational frequency was not immediately obvious the simulation was
repeated but over a longer simulation time. Four different size ratio of 
dimers were studied, both in their 'standard' and 'inverted' orientations. 
The results of this are displayed in Fig.~\ref{fig:rotation_vs_pol}:
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{rotation_vs_pol.png}
	\caption{Rotation frequency vs component phase difference for differently 
		sized dimers. The solid lines represent the rotation rate experienced 
		while the dimer is in its standard orientation, whereas the solid points 
		are for the case where the orientation is inverted. Laser power $= 100\ mW$.}
	\label{fig:rotation_vs_pol}
\end{figure}

A possible explanation of this phenomena was devised by by \cite{Yevick2017}, 
who found that highly focused Gaussian beams could produce second 
order effects in the Rayleigh regime resulting in a photo-kinetic 
force that results in orbital motion about the beam's central axis.
This was verified by \cite{Ruffner2012} who found that by trapping 
a silica bead and recording its position over the course of several 
minutes; while not immediately evident from the trajectory by measuring 
the probability flux of finding a particle in a given position 
revealed a helicity to the sphere's motion which matched the helicity 
of the trapping beam. This would hardly be considered similar to 
the behaviour shown by dimers; firstly because the probability flux 
is an indicator of the sphere's position not its orientation - the 
sphere can be thought to 'orbit' around the central beam axis with a 
radius of orbit on the order of $10^{-9}\ m$. Secondly, while 
\cite{Yevick2017} does compute a rotational frequency, for a single 
sphere this rotation rate is at most $0.01\ Hz$ whereas the lowest 
rotation rate recorded for a dimer ($a_{I}/a_{II}$) was around $2\ Hz$, 
a full two orders of magnitude greater. As such the rotation demonstrated
cannot be explained due to photo-kinetic forces.

This rotation was first noted by Vigilante and co-workers who only 
considered this behaviour for a symmetric dimer \cite{Vigilante2020}; 
to expand upon this work we looked at how the rotation rate changed
with the dimers' size ratio ($a_{I}/a_{II}$) and if the orientation
of the dimer plays any impact on the rotational frequency. By repeating
the same kinds of simulation as used in \ref{fig:rotation_vs_pol} but for
a circularly polarised beam $\phi=90\circ$ it was found that not only is 
the rotation rate dependent on the size of the dimer, but also on its 
orientation and therefore their axial position.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
  \caption{Rotation rate plotted against dimer size ratio while trapped in a
  	circularly polarised beam; a positive rotation rate indicates clockwise 
  	rotation, whereas a negative rotation rate indicates counter-clockwise 
  	rotation. The red line is for the case of a dimer in its 'standard' 
  	orientation. The blue line is for the case when the dimer is in its 
  	'inverted' orientation while trapped above the focus of the beam. And 
  	lastly the green line is for the case when the dimer is in its 
  	'inverted' orientation, but when it is trapped below the focus of the 
  	beam.}
\end{figure}

It is difficult to see from the graph, but the rotation rate never
truly goes down to zero, reaching a minimum of $2\,{\rm Hz}$, which would
imply that a second sphere of radius $200\,{\rm nm}$ is enough to
induce rotational motion. This brings into question what mechanism is
generating the optical torque. We used \textit{mstm} to look at the stokes
parameters from the scattered field from a simple plane wave incident
on a symmetric dimer, the proportion of circularly polarised light is 
minimal compared to the proportion of plane polarised light, which 
indicates that this rotational motion is not due to any inhomogeneity 
in the dimer that might impart angular momentum to the scattered beam - 
as compared to a anisotropic scatterer like vaterite. Therefore the 
rotational motion must be due to

These results are somewhat contradictory to other work with silica dimers
\cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments
have trapped the dimer in an orientation perpendicular to the beam
propagation direction. The rotational motion is attributed to the
dimers asymmetric geometry creating an unbalanced polarisation susceptibility
along its long axis as compared to its short axis; therefore its long
axis is aligned with the polarisation vector and can rotate
freely \cite{Ahn2018}. This however cannot be the case with our
simulations as the dimer rotates about its long axis, meaning there
cannot be an asymmetric axis to align with the beam's polarisation
vector. Furthermore, we see a non-linear increase in the rotational
speed of our dimers with size, the drag torque from the surrounding
fluid is $\propto r^3$ so the expectation is that the rotation
frequency should fall off with increasing size.  This indicates that
the rotational motion is due to the shape asymmetry of the dimer and
not solely due to the beam's angular momentum.  Measurement of this
photo-kinetic force is difficult to achieve due to the fact that
previous analysis was conducted in the Rayleigh regime, where the
polarizability of our dimer can be approximated as:
\begin{align}
  \bold{p}({\bf r},t)
  =
  \alpha_x E_x({\bf r},t)\hat{\bf e}_x
  + \alpha_yE_y({\bf r},t)\hat{\bf e}_y
  + \alpha_zE_z({\bf r},t)\hat{\bf e}_z
\end{align}

@@

\subsection{Gyroscopic Precession using asymmetric dimers}
As mentioned in section~\ref{sec:off-axis} for specificity 
sized dimers there is the potential for non-vertical trapping 
orientations in which the dimer is located in a harmonic 
trap. When trapped by a circularly polarised light these 
dimers exhibit gyroscopic precession. As shown in fig~\ref{fig:gyro} 
the dimer's is rotating about it's long axis (as seen by 
the periodic behaviour of $U_x$ and $U_y$), while also precessing 
around the optical axis of the beam, resulting in a second 
order oscillation in $s_z$.
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped in an 
		off axis orientation with a circuarly polarised beam ($P= 100\ mW$). 
		The far left column depicts the dimer's centre of mass position with 
		time; middle two columns are the $x$, $y$, and $z$ components of the 
		vectors $u_x$ and $u_y$; last column depicts the components of the 
		vector $u_z$ which defines the dimer's orientation.}
	\label{fig:gyro}
\end{figure}

@@

The linear relationship could partly be due to fact that we do not
account for the change in viscous forces with increasing laser power. 
Due to the localised heating effect 
It is far more likely that the rotational frequency reaches a maximum 
value assuming that the bulk fluid can readily absorb the laser.
